Variational Monte Carlo on bosonic systems

### 2018-01-22 project_1/article/article.tex
Add the gradient and Laplacian operators

\section{Theory}
    To model the trapped bosonic gas particles we use the potential
    \begin{align}
        V_{\text{ext}}(\vf{r})
        &=
        \begin{cases}
            \half m\omega^2r^2 & (\text{S}), \\
            \half m \bigl[
                \omega^2(x^2 + y^2) + \omega_z^2z^2
            \bigr] & (\text{E}),
        \end{cases}
    \end{align}
    where we can choose between a spherical (S) or an elliptical (E) harmonic
    trap. The two-body Hamiltonian of the system is given by
    \begin{align}
        H = \sum_{i}^{N}h(\vf{r}_i) + \sum_{i < j}^{N}w(\vf{r}_i, \vf{r}_j),
    \end{align}
    where the single particle one body operator, $h$, is given by
    \begin{align}
        h(\vf{r}_i) = -\frac{\hbar^2}{2m}\nabla_i^2
        + V_{\text{ext}}(\vf{r}_i),
    \end{align}
    and the two-body interaction operator, $w$, is
    \begin{align}
        w(\vf{r}_i, \vf{r}_j)
        = \begin{cases}
            \infty & |\vf{r}_i - \vf{r}_j| \leq a, \\
            0 & |\vf{r}_i - \vf{r}_j| > a,
        \end{cases}
    \end{align}
    where $a$ is the hard sphere of the particle. The trial wavefunction we will
    be looking at is given by
    \begin{align}
        \Psi_T(\vf{r})
        &= \bra{\vf{r}}\ket{\Psi_T}
        = \prod_i^N g(\alpha, \beta, \vf{r}_i)
        \prod_{i < j}^N f(a, \vf{r}_i, \vf{r}_j),
    \end{align}
    where $\alpha$ and $\beta$ are variational parameters and
    \begin{align}
        \vf{r} = (\vf{r}_1, \vf{r}_2, \dots, \vf{r}_N, \alpha, \beta).
    \end{align}
    Here $g$ are the single-particle wavefunctions given by
    \begin{align}
        g(\alpha, \beta, \vf{r}_i)
        = \exp\bigl[
            -\alpha(x_i^2 + y_i^2 + \beta z_i^2)
        \bigr] \equiv \phi(\vf{r}_i),
    \end{align}
    and the correlation wavefunction, $f$, is
    \begin{align}
        f(a, \vf{r}_i, \vf{r}_j)
        &=
        \begin{cases}
            0 & |\vf{r}_i - \vf{r}_j| \leq a, \\
            \Bigl(
                1 - \frac{a}{|\vf{r}_i - \vf{r}_j|}
            \Bigr) & |\vf{r}_i - \vf{r}_j| > a.
        \end{cases}
    \end{align}

@@

        \subsubsection{Non-interacting harmonic oscillator}
            We start by finding an analytical expression for $a = 0$ and $\beta
            = 1$, i.e., a non-interaction harmonic oscillator given by the
            wavefunction
            \begin{align}
                \bra{\vf{r}}\ket{\Psi_T}
                = \prod_i^N \exp\bigl[
                    -\alpha |\vf{r}_i|^2
                \bigr],
            \end{align}
            where $|\vf{r}_i| = r_i$. As $a = 0$ the interaction term,
            $w(\vf{r}_i, \vf{r}_j)$, vanishes and the Hamiltonian is given by
            (in the spherical case)
            \begin{align}
                H &= \sum_i^N h(\vf{r}_i)
                = \sum_i^N \Bigl(
                    -\frac{\hbar^2}{2m}\nabla_i^2
                    + \half m \omega^2 r_i^2.
                \Bigr)
            \end{align}

>>>>

\section{Theory}
    To model the trapped bosonic gas particles we use the potential
    \begin{align}
        V_{\text{ext}}(\vf{r})
        &=
        \begin{cases}
            \half m\omega^2r^2 & (\text{S}), \\
            \half m \bigl[
                \omega^2(x^2 + y^2) + \omega_z^2z^2
            \bigr] & (\text{E}),
        \end{cases}
    \end{align}
    where we can choose between a spherical (S) or an elliptical (E) harmonic
    trap. The two-body Hamiltonian of the system is given by
    \begin{align}
        H = \sum_{i}^{N}h(\vf{r}_i) + \sum_{i < j}^{N}w(\vf{r}_i, \vf{r}_j),
    \end{align}
    where the single particle one body operator, $h$, is given by
    \begin{align}
        h(\vf{r}_i) = -\frac{\hbar^2}{2m}\nabla_i^2
        + V_{\text{ext}}(\vf{r}_i),
    \end{align}
    (we assume equal mass)
    and the two-body interaction operator, $w$, is
    \begin{align}
        w(\vf{r}_i, \vf{r}_j)
        = \begin{cases}
            \infty & |\vf{r}_i - \vf{r}_j| \leq a, \\
            0 & |\vf{r}_i - \vf{r}_j| > a,
        \end{cases}
    \end{align}
    where $a$ is the hard sphere of the particle. The trial wavefunction we will
    be looking at is given by
    \begin{align}
        \Psi_T(\vf{r})
        &= \bra{\vf{r}}\ket{\Psi_T}
        = \prod_i^N g(\alpha, \beta, \vf{r}_i)
        \prod_{i < j}^N f(a, \vf{r}_i, \vf{r}_j),
    \end{align}
    where $\alpha$ and $\beta$ are variational parameters and
    \begin{align}
        \vf{r} = (\vf{r}_1, \vf{r}_2, \dots, \vf{r}_N, \alpha, \beta).
    \end{align}
    Here $g$ are the single-particle wavefunctions given by
    \begin{align}
        g(\alpha, \beta, \vf{r}_i)
        = \exp\bigl[
            -\alpha(x_i^2 + y_i^2 + \beta z_i^2)
        \bigr] \equiv \phi(\vf{r}_i),
    \end{align}
    and the correlation wavefunction, $f$, is
    \begin{align}
        f(a, \vf{r}_i, \vf{r}_j)
        &=
        \begin{cases}
            0 & |\vf{r}_i - \vf{r}_j| \leq a, \\
            \Bigl(
                1 - \frac{a}{|\vf{r}_i - \vf{r}_j|}
            \Bigr) & |\vf{r}_i - \vf{r}_j| > a.
        \end{cases}
    \end{align}

@@

        \subsubsection{Non-interacting harmonic oscillator}
            We start by finding an analytical expression for $a = 0$ and $\beta
            = 1$, i.e., a non-interaction harmonic oscillator given by the
            wavefunction
            \begin{align}
                \bra{\vf{r}}\ket{\Psi_T}
                = \prod_i^N \exp\bigl[
                    -\alpha |\vf{r}_i|^2
                \bigr],
            \end{align}
            where $|\vf{r}_i| = r_i$. As $a = 0$ the interaction term,
            $w(\vf{r}_i, \vf{r}_j)$, vanishes and the Hamiltonian is given by
            (in the spherical case)
            \begin{align}
                H &= \sum_i^N h(\vf{r}_i)
                = \sum_i^N \Bigl(
                    -\frac{\hbar^2}{2m}\nabla_i^2
                    + \half m \omega^2 r_i^2
                \Bigr).
            \end{align}
            Working in spherical coordinates the gradient and the Laplace
            operator (applied to a function) becomes
            \begin{align}
                \nabla_if
                &= \vf{e}_r\tpl{r}f
                + \vf{e}_{\theta}\frac{1}{r}\tpl{\theta}f
                + \vf{e}_{\phi}\frac{1}{r\sin(\theta)}\tpl{\phi}f, \\
                \nabla_i^2f
                &= \frac{1}{r^2}\tpl{r}
                \bigl[
                    r^2\tpl{r}f
                \bigr]
                + \frac{1}{r^2\sin(\theta)}\tpl{\theta}
                \bigl[
                    \sin(\theta)\tpl{\theta}f
                \bigr] \nonumber\\
                &\qquad
                + \frac{1}{r^2\sin^2(\theta)}\tpl{\phi}^2f,
            \end{align}
            where $\vf{e}_r$ signifies the unit vector in the radial direction
            and we use the shorthand notation
            \begin{align}
                \tpl{r}f \equiv \dpd[]{f}{r}.
            \end{align}
